Eliminating object prior-bias from sparse-projection tomographic reconstructions

### 2021-10-15 paper/post_tci.tex
some textual changes.

 \begin{figure}[t]
\centering
	\includegraphics[width=0.45\textwidth]{../images/story/post_TCI/story.png}
        \caption{Illustration of the use of priors in reconstruction
          as the number of views reduces. For all cases, the ground
          truth (260$\times$260) appears in the top left.  For
          extremely large number of views (800 views), the standard
          FBP reconstruction (first column) is of very good quality
          (Structural Similarity Index Metric, SSIM=0.92). As the
          number of views become limited (100 views), reconstruction
          using the Total Variation regularization prior (TV, middle
          column) is better both visually and quantitatively than FBP
          (SSIM of 0.96 vs 0.82). If the number of views is
          drastically reduced (20 views), the presence of both TV
          prior and object-prior together with the spatially-varying
          technique of this paper improves the reconstruction
          (SSIM=0.91) when compared to presence of TV-only prior
          (SSIM=0.87) or no prior (SSIM=0.64).}
 \label{fig:story}
 \end{figure}

  When the number of views is reduced significantly further (as is discussed in this work), a
 regularization prior alone is not sufficient.  In such cases
 additional information (`object-prior') specific to the current
 object being scanned (called the `test' henceforth) is useful in
 further improving the reconstruction.  While this has been done in
 the context of dynamic CT-scans (discussed more in
 Section~\ref{sec:related}), in this paper we use a set $S$ of
 previous scans of the same object. This is relevant in a longitudinal
 context: the acquisition of sequential CT scans of the same subject
 in order to track time-evolving changes within the subject's
 interior. 
 \subsection{Difficulties}

 With the use of an object-prior, a \emph{new} challenge emerges. The prior
 set $S$ may potentially overwhelm the necessary details, and when
 several prior scans are available, finding the best one is an
 issue. We therefore devise an algorithm that estimates the location
 and magnitude of new changes in the (unknown) test. As we show in
 Section~\ref{sec:method_spatially_varying_prior}, this prevents the prior from adversely affecting the reconstruction of new
 regions in the test. We refer to this method as a
 \textit{spatially-varying} prior-based reconstruction routine, that
 still uses \textit{all} the previous scans in the set $S$, without
 using any of the prior scans \textit{as is}, or endeavoring to choose
 one ``right'' prior. In this way, this work is novel and compared to
 earlier work. As seen in Fig.~\ref{fig:story} (last column, 20
 views) our method uses the set $S$ to yield a better reconstruction
 than the regularization-only, or the backprojection-only methods.

@@

\textit{Remark:} Unlike the potato, the test was designed as one with
the absence of a feature.  Although mathematically one is the
conjugate of the other, in practice, experimental methods may have
difficulties which explains the choice of our test. Notice also that
the structure of the Okra is intricate as compared to Potato, and this
brings with it a stronger challenge.

@@

$\lambda_1$ and  $\lambda_2$ are tunable weights given to the
TV-prior and object-prior terms respectively. Their selection is discussed in
Section.~\ref{sec:tuning_parameters}. The unknowns
$\boldsymbol{x}$ and $\boldsymbol{\alpha}$ are solved by alternately
minimizing $J_{\boldsymbol{\alpha}}(\boldsymbol{x})$ using a fixed
$\boldsymbol{\alpha}$, and $J_{\boldsymbol{x}}(\boldsymbol{\alpha})$
using the resultant $\boldsymbol{x}$.
$J_{\boldsymbol{\alpha}}(\boldsymbol{x})$ is solved for using an
optimal first-order method for large-scale TV regularization presented
in~\cite{TVReg} and available in~\cite{TVReg-lib}.

@@

\subsection{Tuning of parameters}
\label{sec:tuning_parameters}
$\lambda_1$, $\lambda_2$ and $k$ are the three hyper-parameters that
need to be chosen carefully for optimal reconstruction. In our
experiments, $\lambda_1$ was tuned in such a way as to maximize the
SSIM of the complete reconstructed test volume using the TV-only
regularizer.  In particular, the RoI was not considered.  Now, this
value was retained for our spatially-varying prior-based
reconstruction as well.  In principle, one could have jointly
optimized both $\lambda_1$ and $\lambda_2$ to get even better results
compared to the one we have in Sections~\ref{sec:results_with_imaging_model}, and~\ref{sec:results_without_imaging_model}.  The value of
$\lambda_2$ largely depends on the amount of artefacts one aims to
remove by using object-prior at the cost of their dominance in the new
regions. This value was chosen to lie between $0-3$ for our datasets
(having the fixed $\lambda_1$).

@@

\subsubsection{Symbolic}
Let
\begin{itemize}
\item $N$ denote the number of voxels of test volume,
\item $L$ denote the number of templates available,
\item $M$ denote the number of pilot reconstruction algorithms used for the computation of weights map, and
\item $F$ denote the time taken for optimization.
  $F$ consists of time taken for alternate minimization of Eq.~\ref{eq:spatially_varying_prior}, which in turn consists of time taken for repeated cycles of a closed-form update of $\boldsymbol{\alpha}$ (Eq.~\ref{eq:compute_alpha}) and an iterative optimization of  $\boldsymbol{x}$ (Eq.~\ref{eq:compute_x}). 
\end{itemize}
We note that in practice $L$ is independent of $N$ and is likely to be
$O(1)$ (e.g., about 12 scans in a year for disease prognosis).
Likewise $M$ is also $O(1)$. The speed of the iterative optimization
is dependent on the number of imaging views, and is profoundly
impacted by the actual numerical optimization technique used.
Therefore, the time computational complexity of our method is sum of
the following
\begin{enumerate}
\item $O(NL^2+L^3)$ for creating eigenspaces, which is, in practice linear, i.e., $O(N)$.
\item $O(NLM)$ for deriving weights map, which is, in practice linear, i.e., $O(N)$.
\item and the time $F$  taken for iterative minimization, which as mentioned above is not the key contribution of this paper. 
\end{enumerate}
In short, the additional burden of the key idea of this paper viz.,
generating weights map is linear in the number of voxels, although
admittedly the burden $F$ of the alternate minimization procedure
drives up the cost (as can be seen in Table~\ref{table:compute_time}).

>>>>

 \begin{figure}[t]
\centering
	\includegraphics[width=0.45\textwidth]{../images/story/post_TCI/story.png}
        \caption{Illustration of the use of priors in reconstruction
          as the number of views reduces. For all cases, the ground
          truth (260$\times$260) appears in the top left.  For
          extremely large number of views (800 views), the standard
          FBP reconstruction (first column) is of very good quality
          (Structural Similarity Index Metric, SSIM=0.92, ideal SSIM=1.00). As the
          number of views become limited (100 views), reconstruction
          using the Total Variation regularization prior (TV, middle
          column) is better both visually and quantitatively than FBP
          (SSIM of 0.96 vs 0.82). If the number of views is
          drastically reduced (20 views), the presence of both TV
          prior and object-prior together with the spatially-varying
          technique of this paper improves the reconstruction
          (SSIM=0.91) when compared to presence of TV-only prior
          (SSIM=0.87) or no prior (SSIM=0.64).
          % , an improvement by $4\%$ and $27\%$ respectively (since the ideal SSIM=1)
}
 \label{fig:story}
 \end{figure}

 When the number of views is reduced significantly further (as
 discussed in this work), a regularization prior alone is not
 sufficient.  In such cases additional information (`object-prior')
 specific to the current object being scanned (called the `test'
 henceforth) is useful in further improving the reconstruction.  While
 this has been done in the context of dynamic CT-scans (discussed more
 in Section~\ref{sec:related}), in this paper we use a set $S$ of
 previous scans of the same object. This is relevant in a longitudinal
 context: the acquisition of sequential CT scans of the same subject
 in order to track time-evolving changes within the subject's
 interior.
 \subsection{Difficulties}

 With the use of an object-prior, a \emph{new} challenge emerges. The prior
 set $S$ may potentially overwhelm the necessary details, and when
 several prior scans are available, finding the best one is an
 issue. We therefore devise an algorithm that estimates the location
 and magnitude of new changes in the (unknown) test. As we show in
 Section~\ref{sec:method_spatially_varying_prior}, this prevents the prior from adversely affecting the reconstruction of new
 regions in the test. We refer to this method as a
 \textit{spatially-varying} prior-based reconstruction routine, that
 still uses \textit{all} the previous scans in the set $S$, without
 using any of the prior scans \textit{as is}, or endeavoring to choose
 one ``right'' prior. In this way, this work is novel and to be
 contrasted with 
 earlier work. As seen in Fig.~\ref{fig:story} (last column, 20
 views) our method uses the set $S$ to yield a better reconstruction
 than the regularization-only, or the backprojection-only methods.

@@

\textit{Remark:} Unlike the potato, the test was designed as one with
the absence of a feature.  Although mathematically one is the
conjugate of the other, in practice, experimental methods may have
difficulties which explains the choice of our test. Notice also that
the structure of the Okra is intricate as compared to Potato, and this
brings with it a different challenge.

@@

$\lambda_1$ and $\lambda_2$ are tunable weights given to the TV-prior
and object-prior terms respectively and the specification of these are
discussed in Section.~\ref{sec:tuning_parameters}. The unknowns
$\boldsymbol{x}$ and $\boldsymbol{\alpha}$ are solved by alternately
minimizing $J_{\boldsymbol{\alpha}}(\boldsymbol{x})$ using a fixed
$\boldsymbol{\alpha}$, and $J_{\boldsymbol{x}}(\boldsymbol{\alpha})$
using the resultant $\boldsymbol{x}$.
$J_{\boldsymbol{\alpha}}(\boldsymbol{x})$ is solved for using an
optimal first-order method for large-scale TV regularization presented
in~\cite{TVReg} and available in~\cite{TVReg-lib}.

@@

\subsection{Tuning of parameters}
\label{sec:tuning_parameters}
$\lambda_1$, $\lambda_2$ and $k$ are the three hyper-parameters that
need to be chosen carefully for optimal reconstruction. In our
experiments, $\lambda_1$ was tuned in such a way as to maximize the
SSIM of the complete reconstructed test volume using the TV-only
regularizer.  In particular, regions of interest (RoI) are obviously
not known in advance, and thus not considered for the purpose of
tuning.  Further, this value was retained for our spatially-varying
prior-based reconstruction as well, that is, the value optimized for
the TV-only regularize is maintained for the method proposed in this
paper.  In principle, one could have jointly optimized both
$\lambda_1$ and $\lambda_2$ to get even better results compared to the
one we have in Sections~\ref{sec:results_with_imaging_model},
and~\ref{sec:results_without_imaging_model}.  The value of $\lambda_2$
largely depends on the amount of artefacts one aims to remove by using
object-prior at the cost of their dominance in the new regions. This
value was empirically chosen to lie between $0--3$ for our datasets
(having the fixed $\lambda_1$).

@@

\subsubsection{Symbolic}
Let
\begin{itemize}
\item $N$ denote the number of voxels of test volume,
\item $L$ denote the number of templates available,
\item $M$ denote the number of pilot reconstruction algorithms used for the computation of weights map, and
\item $F$ denote the time taken for optimization.
  $F$ consists of time taken for alternate minimization of Eq.~\ref{eq:spatially_varying_prior}, which in turn consists of time taken for repeated cycles of a closed-form update of $\boldsymbol{\alpha}$ (Eq.~\ref{eq:compute_alpha}) and an iterative optimization of  $\boldsymbol{x}$ (Eq.~\ref{eq:compute_x}). 
\end{itemize}
We note that in practice $L$ is independent of $N$ and is likely to be
$O(1)$ (e.g., about 12 scans in a year for disease prognosis).
Likewise $M$ is also $O(1)$. The speed of the iterative optimization
is dependent on the number of imaging views, and is profoundly
impacted by the actual numerical optimization technique used.
Therefore, the time computational complexity of our method is sum of
the following
\begin{enumerate}
\item $O(NL^2+L^3)$ for creating eigenspaces, which is, in practice linear, i.e., $O(N)$.
\item $O(NLM)$ for deriving weights map, which is, in practice linear, i.e., $O(N)$.
\item and the time $F$  taken for iterative minimization, which as mentioned above is not the key contribution of this paper. 
\end{enumerate}
In short, the additional burden of the key idea of this paper viz.,
generating weights map is linear in the number of voxels, although
admittedly the burden $F$ of the alternate minimization procedure
drives up the cost (as can be seen in
Table~\ref{table:compute_time}). The supplementary material describes
a method to avoid this cost if absolutely necessary.
